\documentclass{article}

\usepackage{amsmath}

\begin{document}



\section{Boid Behavior}


Each boid maintains a position vector, $\vec{p}$, and a velocity
vector, $\vec{v}$.  At each time step the boid calculates an
acceleration vector,$\vec{a}$ based on local rules described below.
The new velocity and positions vectors are calculated as:
\begin{align*}
  \vec{v}_{t+1} &= \vec{v}_t + \vec{a}_t \Delta t \\
  \vec{p}_{t+1} &= \vec{p}_t + \vec{v}_{t+1} \Delta t
\end{align*}

\paragraph{Creating the acceleration vector}
The acceleration vector is calculated from three vectors:
\begin{enumerate}
\item the difference between the mean position of the boid's local
  neighbors (radius for local neighborhood is a parameter) and the
  boid's current position,
\item the difference between the mean velocity of the boid's local
  neighbors and the boid's current velocity, and,
\item a weighted sum of differences between close neighbors' positions
  (controlled by a difference radius parameter) and boid's current
  position as well as differences from stationary objects.
\end{enumerate}

Vector 1 represents the desire to be in the middle of the flock,
vector 2 represents the desire to fly in the same direction and vector
3 encodes obstacle avoidance.  Each of these desires is given a user
controlled weight and the acceleration vector is constructed using one
of two allocation techniques.  The length of acceleration vector is
limited by a global constant.

\paragraph{mean allocation} The three component vectors are weighted,
summed, and rescaled to the maximum acceleration.

\paragraph{priority allocation} The three component vectors are weighted, ranked
and added to the acceleration vector in amounts that respect the
maximum acceleration.  Vector 3 is of highest priority so if after
weighting its length exceeds the maximum acceleration it is rescaled
and used as the acceleration.  If the weighted highest priority vector
does not exceed the maximum acceleration we add as much of the
weighted second
priority (vector 2) to the weighted highest priority vector as
follows.

Let
$\vec{x}$ be the weighted highest priority vector and $\vec{y}$ the
weighted second priority vector.  Set

\begin{equation*}
  \alpha = \frac{-2 (\vec{x} \cdot \vec{y}) + \sqrt{(2 \vec{x} \cdot
      \vec{y})^2 + 4 (\vec{y}\cdot \vec{y}) (m^2 -
      \vec{x}\cdot\vec{x})}}{2( \vec{y}\cdot\vec{y})}
\end{equation*}
where $m$ is the maximum acceleration and the dot product has the
usual definition: $\vec{x}\cdot\vec{y} = \langle x_1,x_2,x_3 \rangle
\cdot \langle y_1,y_2,y_3 \rangle = x_1y_1 + x_2y_2 + x_3y_3$.




If $\alpha > 1$ set the current acceleration vector to $\vec{z} =
\vec{x} + \vec{y}$ and continue by adding in the maximum amount of the
weighted third priority in the same fashion.  If $\alpha \leq 1$ the
acceleration is $\vec{z} = \vec{x} + \alpha \vec{y}$.


\end{document}